Evaluating infectious disease forecasts with allocation scoring rules

### 2023-07-12 alloscore_manuscript/allocation.tex
Begin allocation solution algorithm description

@@

We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to the states.

>>>>

\subsection{Solving the Allocation Problem}

As the solution to a constrained optimization problem, the optimal allocation $\bf{x}^F$ according to a forecast $F$ can be described formally as the first component of a sationary point $(\bf{x}^F, \lambda^{\star})$ for the Lagrangian 
\begin{align}
\mathcal{L}(\mathbf{x},\lambda; K, L) = \sum_{i=1}^{N} \mathbb{E}_{F_i}[s_A(x_i, Y_i; L)] + \lambda\left(\sum_{i=1}^{N} x_i - K\right).
\end{align}
That is, optimality requires that both
\begin{align}
0 = \nabla_{x} \mathcal{L}(\mathbf{x}^F, \lambda) &= \sum_{i=1}^{N} E_F \left[\frac{d}{dx_i}(Y_i - x_i^F)_{+}\right] + \lambda \\
& = \sum_{i=1}^{N} E_F \left[-\mathbf{1}(Y_i \geq x_i^F)\right] + \lambda  = \sum_{i=1}^{N} F_i(x_i^F)-1 + \lambda  \label{eqn:x_stationary}\\
\text{ and} \\
0 = \nabla_{\lambda} \mathcal{L}(\mathbf{x}^F, \lambda^{\star}) &= \sum_{i=1}^{N} x_i^F - K\\
& = \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K \quad \text{ (by \eqref{eqn:x_stationary})}. \label{eqn:lambda_stationary}
\end{align}
These equations give formulae $x_i^F = F_i^{-1}(1-\lambda^{\star})$, but in many situations of interest for us the $F_i^{-1}$ are not single-valued, and even 
when they are the solution $\lambda^{\star}$ of \eqref{eqn:lambda_stationary} can be obtained analytically only in special cases (but see \ref{sec:methods.ls.example}). We therefore approximate allocations using a binary search for $\lambda^{\star}$ beginning with the interval $\lambda \in [0,1]$ and 
checking at each step the sign of the inequality that results from substituting the midpoint $\lambda_{\tau}$ into \eqref{eqn:lambda_stationary} for endpoints of the sets $F_i^{-1}(1-\lambda_{\tau})$.  If $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) > K$, the quantiles for $1-\lambda_{\tau}$ exceed our resource constraint so we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.  Similarly, if $\sum_{i} F_{i}^{-1}(1-\lambda^{\star}) < K$, the quantiles for 
$1-\lambda_{\tau}$ do not make ful use of our resource and we set $\lambda_{\tau}$ as the lower endpoint of the $\tau + 1$ search interval.  This process halves the search interval width at each step and we continue until this width falls below a given tolerance level $\varepsilon_{\tau}$.

@@

We illustrate with an application to hospital admissions in the U.S., considering the problem of allocation of a limited supply of medical resources to {}the states.
